---
title: "re_ml_01"
author: "Noah Suter"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: yes
---
# AGDS Report

## Report Exercise 5

### Preparing the data

```{r results='hide'}
list.files('../data') #listing the files to find the right data
```


```{r results='hide', message=FALSE}
half_hourly_fluxes <- readr::read_csv("../data/df_for_stepwise_regression.csv") # reading the data into R
```


```{r}
visdat::vis_miss(half_hourly_fluxes) #checking for missing values
```

```{r results='hide'}
summary(half_hourly_fluxes) #looking at the exact number of missing values (NA) per row
```

```{r message=FALSE, warning=FALSE}
half_hourly_fluxes_clean <- stats::na.omit(half_hourly_fluxes) #deleting all the rows with missing values
```


```{r message=FALSE, warning=FALSE}
library(ggplot2)
# Data cleaning: no obvious bad data noticable
# as there is no long tail, no further target engineering
half_hourly_fluxes_clean |> 
  ggplot(aes(x = GPP_NT_VUT_REF, y = ..count..)) + 
  geom_histogram()
```

## Task 1 - Comparison of the linear regression and KNN models


```{r message=FALSE, warning=FALSE}
source("../functions/large_files_ex_5.R") #loading all the large code chunks form another .R file
```

```{r warning=FALSE, message=FALSE}
eval_model(mod = mod_lm, df_train = half_hourly_fluxes_train, df_test = half_hourly_fluxes_test) #running the lm model with test and train data
```
```{r warning=FALSE, message=FALSE}
# KNN model
eval_model(mod = mod_knn, df_train = half_hourly_fluxes_train, df_test = half_hourly_fluxes_test) #running the knn model with train and test data
```

###Interpretation

**Question 1: Why is the difference between the evaluation and the test set larger for the KNN model than for the linear regression model?**

The bigger difference of the KNN model suggests that it might be overfitting. This can be expected as the KNN models are prone to overfitting.
More exaclty the R2 indicates a drop in performance as the R2 decreases from the training to the test set.
The RMSE increases which indicates a higher predicted error for the test set.

**Question 2: Why does the evaluation on the test set indicate a better model performance of the KNN model than the linear regression model?**

As the KNN model has a higher R2 and a lower RMSE in comparison to the lm model. (which indicates a better model performance) But it is difficult to say which model is better in the end as the difference is not that big.

**Question 3: How would you position the KNN and the linear regression model along the spectrum of the bias-variance-trade-off?**

typically the KNN has a higher variance and therefore a lower bias.
The lm typically is lower variance and therefore higher bias

And the same can be seen within the results. Because the larger difference for the knn model indicates a rather higher variance

the lm has a smaller difference and therefore a rather higher bias and a lower variance.


### Visualisation of the temporal variations of observed and modelled GPP for both models

```{r}
library(ggplot2)
# create a dataframe with the dates and observed GPP
observed <- data.frame(Date = half_hourly_fluxes_clean$TIMESTAMP, GPP = half_hourly_fluxes_clean$GPP_NT_VUT_REF)
# add columns with modelled GPP for lm and KNN models
observed$lm <- predict(mod_lm, newdata = half_hourly_fluxes_clean)
observed$knn <- predict(mod_knn, newdata = half_hourly_fluxes_clean)
# create a line plot that shows observed and modelled GPP for linear regression models
ggplot(observed, aes(x = Date)) +
  geom_line(aes(y = GPP, color = "Observed")) +
  geom_line(aes(y = lm, color = "Linear Regression")) +
  labs(x = "Date", y = "GPP", color = "Model") +
  scale_color_manual(values=c("green", "blue")) +
  ggtitle("Observed and Modelled GPP for Linear Regression")
```
```{r}
# create a line plot that shows observed and modelled GPP for KNN models
ggplot(observed, aes(x = Date)) +
  geom_line(aes(y = GPP, color = "Observed")) +
  geom_line(aes(y = knn, color = "KNN",)) +
  labs(x = "Date", y = "GPP", color = "Model") +
  scale_color_manual(values=c("green", "blue")) +
  ggtitle("Observed and Modelled GPP for KNN")
```



## Task 2 - The role of k

### Hypothesis

if k is going towards 1, the models becomes more complex and flexible. This can be explained by the fact, that the prediction for a new data point is based on the closest neighbor. In terms of overfitting this can turn into a problem.if we then guess the development for R squared we can expect R squared to get bigger if closer to one for the training set. At the same time the problem of overfitting is that the model is getting less generalizability. Therefore it can be expected that the r squared for the test set is decreasing if closer to one.
The MAE is getting lower in the overfitted data. Therefore in the training set one can expect a lower MAE if closer to one (as it is more overfitted if closer to one). Yet the same problem with the generalizability accurs doe to the overfitting. Therefore the MAE would get bigger for the test set if closer to one.

In the direction of k = N the problem of underfitting can accur. Therefore the values will develop the other way around for higher k.

In general a bias-variance trade-off is happening. if k is small, the model has low bias but high variance. If k is large the model has high bias but low variance. For a good solutions a compromise needs to be found.

### Hypothesis put to test with a code

```{r message=FALSE, warning=FALSE}
# Evaluate KNN model for different values of k
k_values <- c(1:30)
mae_values <- sapply(k_values, function(k) get_mae(k)$mae_test)
mae_train_values <- sapply(k_values, function(k) get_mae(k)$mae_train)
```

```{r}
# showing model generalisability  
df <- data.frame(k = k_values, MAE_test = mae_values, MAE_train = mae_train_values)
ggplot(df, aes(x = k)) + 
  geom_line(aes(y = MAE_test, color = "Test")) + 
  geom_line(aes(y = MAE_train, color = "Train")) +
  geom_point(aes(y = MAE_test)) + 
  geom_point(aes(y = MAE_train)) + 
  xlab("k") + 
  ylab("MAE") +
  scale_color_manual(name = "Data set", values = c("red", "blue"))
```




```{r warning=FALSE}
get_mae(k=15) #getting the mae for the train and test data with k=15 as example
```



### Code for trying to find an optimal k

```{r}
# Determinen the optimal k value
optimal_k <- k_values[which.min(mae_values)]

# Print the corresponding results
cat("Optimal k value for generalizability:", optimal_k, "\n")
cat("MAE on test set for optimal k value:", mae_values[which.min(mae_values)], "\n")
cat("MAE on training set for optimal k value:", mae_train_values[which.min(mae_values)], "\n")

```


