---
title: "report"
author: "Noah Suter"
date: "2023-02-27"
output: html_document
---

# AGDS Report

## Report Exercise 1

### Task 1


```{r results='hide'}
list.files('../data') #listing the files to find the right data
```

```{r results='hide', message=FALSE}
database_s1<- readr::read_csv("../data/data_csv_exercise_1.csv") # reading the .csv into R
```

```{r}
visdat::vis_miss(database_s1) #checking for missing values
```

The missing values all occur within the column of the sample date. Thus, there are experiments in which the date of execution is not given. This does matter in the way that the experiments are not being separated, even though they would have been in different years (Important for the task 2 where the data is grouped by the experiment and the sample date). But the rest of the data is not missing and "only" the date of the execution is not given all the time. 


### Task 2

```{r results='hide'}
sample_data_RR <- dplyr::mutate(database_s1, log_response_ratio = log((`increased CO2`/`ambient CO2`))) #calculate the RR and add as a new column
```

```{r results='hide', message=FALSE} 
sample_date_yearly <-  sample_data_RR|> #create new table
  dplyr::group_by(Experiment, `Sample date (year)`) |> #aggregate by experiment with subgroup sample date (in years)
  dplyr::summarise(`log_response_ratio_mean` = mean(`log_response_ratio`)) #calculating the mean of RR within the defined groups
```

```{r}
knitr::kable(sample_date_yearly[1:10,])
#present the results as knitr::kable()
```

### Task 3
```{r}
phase_1 <- dplyr::filter(sample_data_RR, `Time (years)` <3) |>
  dplyr::summarise(`log_response_ratio_mean` = mean(`log_response_ratio`)) #aggregate the RR of phase 1
```

```{r}

phase_2 <- dplyr::filter(sample_data_RR, `Time (years)` >=3 & `Time (years)`<6) |>
  dplyr::summarise(`log_response_ratio_mean` = mean(`log_response_ratio`)) #aggregate the RR of phase 2
```

```{r}
phase_3 <- dplyr::filter(sample_data_RR, `Time (years)` >=6) |>
  dplyr::summarise(`log_response_ratio_mean` = mean(`log_response_ratio`)) #aggregate the RR of phase 3
```

```{r}
all_phases <- rbind(phase_1, phase_2, phase_3) #bind all 3 phase tables to one table
all_phases <- dplyr::mutate(all_phases,`Phase` = c("early phase (<3 years since start)", "mid phase (3-6 years since start", "late phase (>6 years since start)" )) #add a column with the phase description
```

```{r}
all_phases <- dplyr::select(all_phases,`Phase`, everything()) #moving the phase column to position one
```

```{r}
knitr::kable(all_phases) #present the results as knitr::table
```









